Towards a Complete Model of the Negative Selection Process in Humans

### 2023-06-07 manuscript/manuscript.tex
new supp, ms stuff

\begin{abstract}

\end{abstract}

@@

While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
prediction for the draft-effective population size $N_d$, closed-form
expressions for the rate of the ratchet $R$ have generally been hard to find
\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
\parencite{Santiago2016-mu} is that the ratchet rate under draft is given by
the probability of fixation $p_f(N_d, s)$
\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
population size, e.g. $R = N U p_f(N_d, s)$. Given this equation for the
ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
system of two non-linear equations that can be solved numerically for $N_d$ and
$R$ for each segment,

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{figures/figure_1.pdf}
  \caption{Theory compared to simulation results. (A) The predicted reduction
      factor under classic B theory (dark gray line) and the diploid SC16 model
      (colored lines corresponding to mutation rate) compared to average
      reduction across XXX simulation replicates (points). (B) The predicted
      ratchet rate under the SC16 model scaled by mutation rate (colored lines)
      compared to the ratchet estimated from simulation (points). When $2Ns>1$,
      the ratchet rate is near zero. (C) The genic variance from simulations
      (points) against the predicted variance under the SC16 model (colored
      lines). As the ratchet begins to click, the genic variance is decreased
      from the level expected under strong BGS (dashed line). (D, bottom) The
      mean squared error (MSE) between whole-chromosome simulations and
      predicted classic B (dots), new B' (solid), and locally-rescaled B'
      (dashed) for different mutation rates (colors). The dashed horizontal
      line is the approximate theoretic minimum MSE. (D, top) The build up of
      negative linkage disequilibria around $2Ns=1$ in whole-chromosome
      simulations. (E) The average B map from 100 chromosome 10 simulation
  replicates (gray) against different predictions, for parameters that
correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. The chromosome shows 
the density of conserved sites and recombination map used in simulations. }
  \label{fig:figure-1}
\end{figure}

@@

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is greater
when mutation rates are higher (Supplementary Figure XXX). This is consistent
our view that the equilibrium fitness variance under the SC16 model is the
additive \emph{genic} fitness variance, which differs from the additive genetic
variance by the sum of linkage disequilibria between selected sites,
$\delta_{LD} = s^2 \sum_{i\nej} D_{i,k}$.

@@

@@

@@

>>>>

\begin{abstract} 
% TODO: emphasize connection to nearly neutral theory
\end{abstract}

@@

While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
prediction for the draft-effective population size $N_d$, closed-form
expressions for the rate of the ratchet $R$ have generally been hard to find
\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
\textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
by the probability of fixation $p_f(N_d, s)$
\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
population size, e.g. $R = N U p_f(N_d, s)$. Given this equation for the
ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
system of two non-linear equations that can be solved numerically for $N_d$ and
$R$ for each segment,

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Theory
        compared to simulation results. (A) The predicted reduction factor
        under classic B theory (dark gray line) and the diploid SC16 model
        (colored lines corresponding to mutation rate) compared to average
        reduction across XXX simulation replicates (points). (B) The predicted
        ratchet rate under the SC16 model scaled by mutation rate (colored
        lines) compared to the ratchet estimated from simulation (points). When
        $2Ns>1$, the ratchet rate is near zero. (C) The genic variance from
        simulations (points) against the predicted variance under the SC16
        model (colored lines). As the ratchet begins to click, the genic
        variance is decreased from the level expected under strong BGS (dashed
        line). (D, bottom) The mean squared error (MSE) between
        whole-chromosome simulations and predicted classic B (dots), new B'
        (solid), and locally-rescaled B' (dashed) for different mutation rates
        (colors). The dashed horizontal line is the approximate theoretic
        minimum MSE. (D, top) The build up of negative linkage disequilibria
        around $2Ns=1$ in whole-chromosome simulations in bottom panel. (E) The
        average B map from 100 chromosome 10 simulation replicates (gray)
    against different predictions, for parameters that correspond to $2Ns < 1$,
$2Ns = 1$, and $2Ns > 1$. The chromosome shows the density of conserved sites
and recombination map used in simulations. }
  \label{fig:figure-1}
\end{figure}

@@

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
when mutation rates are higher (Supplementary Figure XXX). As $Ns \to 0$, the
variance in LD inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}.
Overall, the build up of negative LD is consistent our view that the
equilibrium fitness variance modeled by the SC16 theory is the additive
\emph{genic} fitness variance, which differs from the additive genetic variance
by the sum of linkage disequilibria between selected sites (i.e. $\delta_{LD} =
s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$ is the LD between sites $i$ and
$j$). However, according to theory, the reductions in diversity should be
determined by levels of additive genetic fitness variance that include the
contribution of LD (Supplementary Materials Section XXX).

@@

Next, we developed a composite-likelihood method to estimate pairwise diversity
in the absence of linked selection $\pi_0$, the deleterious mutation rate
($\mu_D$), and distribution of fitness effects ($\mathbf{W}$) under negative
selection. Because the focus of our work is to fix deficiencies in the classic
BGS model, our current implementation does not consider reductions in diversity
due to recurrent hitchhiking or other positive selection processes. Since
previous work has established that purifying selection is the dominant mode of
linked selection in humans and hitchhiking plays a limited role
\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
test our method.

Given that previous methods used classic BGS theory to fit genome-wide patterns
of human diversity, we were curious whether their qualitative findings would
change once weak selection was more accurately predicted. In particular, the
U-shaped relationship between $B'$ and the strength of selection suggests that
a given reduction factor could be explained by two different selection
coefficients.

@@

Like other methods, our method divides the putatively conserved basepairs up

Generally, when the putatively conserved region is larger, there is a tradeoff
in two dimensions. First, more bases are conserved according to the same DFE,
so the total mutational load increases. This causes the tradeoff between
mutation rate estimate and CADD fraction we observe across both our B' model
and the classic B when using sparse tracks (Figure 2XXX).

We find that mutation rate estimates are fairly sensitive to model choice
(Figure 2XXX).

Since B' has a U-shaped relationship with the selection coefficient, we
hypothesized that our model may reveal true identifiability issues in
estimating the DFE that do not occur under the classic B model. We find that
this is not the case for our models fit at the megabase-scale. The DFE
estimates for our CADD6 and CADD8 models are consistent (Figure XXX), and both
produce similar mutation rates.

While we do not comprehensively fit all of our model to other spatial scales,
we

However, while a given reduction level may be explained by either a weak or
strong selection coefficient, the spatial extent of these reductions caused by
weak selection is much narrower than under strong selection.

We see evidence this is the case in several of our models, where
genome-wide diversity fits equally well between models with mutation rates

Indeed, we see this by comparing parameter estimates of the CADD6
sparse-track model using classic BGS B and our B' method.

(which mimics the best-fitting model of
\cite{Murphy2022-sj})

(see also
Appendix 1, Figure 26A of \cite{Murphy2022-sj})

correlation between s and feature

Overall, we find evidence that negative selection models are very sensitive to
the input track. Because inference based on our B' method accommodates weaker
selection, the primary models we fit here more robust to misspecification of
the regions

@@

We calculate the mean squared error (MSE) and relative error between
whole-chromosome simulations and the predicted B and B' maps by averaging
branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
noise in estimating the expected reduction map from simulations. For $n=100$
replicates, the sampling noise is minimal relative to the evolutionary noise in
the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
\approx \nicefrac{2}{9n}$.
